\section{Field Extensions}

We build up the theory of field extensions following Morandi~\cite{Morandi1996}.

The study of field theory is primarily the study of field extensions. The classical problems of ruler-and-compass constructions and the solvability of polynomial equations by radicals were settled by analyzing appropriate field extensions. The central idea is to associate a group---the \emph{Galois group}---to a field extension, converting field-theoretic questions into group-theoretic ones. Since the Galois group of a finite extension is finite, the rich structure theory of finite groups becomes available. In this section we build the foundational vocabulary: degrees, algebraic elements, minimal polynomials, and the key structural results about simple and algebraic extensions.

%------------------------------------------------------------------
\subsection{Degrees and Generated Subfields}
%------------------------------------------------------------------
Let $F\subseteq K$ are fields. 
Then $K$ is called a \emph{field extension} of $F$. 
We denote it by $K/F$. 
Here $F$ is called the \emph{base field}. 
We can think of $K$ as a vector space over $F$ where the addition of two vectors follows the addition operation of $K$ and the multiplication of any $\alpha\in K$ and $c\in F$ follows the multiplication operation of $K$, $c\cdot \alpha=c\alpha$. 
Then we denote the dimension of $K$ over $F$ by  $$[K\colon F]\coloneqq\dim_F(K)$$This dimension is called the \emph{degree of field extension} of $F$ or simply \emph{degree of} $[K\colon F]$. 
If $[K\colon F]<\infty$ we call it by a special name, \emph{finite extension}. 
Finite extensions of finite fields are of very important to us and this report we will work on finite fields and their finite extensions. 
Let $\alpha\in K$. 
Then we denote $F(\alpha)$ to be the field generated by $F$ and $\alpha$ following the multiplication and addition operations of $K$. 
We will define what it means to be the field generated by $F$ and $\alpha$ in \defref{def:generated-subfields}.
It is very easy to sea $F(\alpha)$ is also a field, and it contains $F$ but also contained in $K$. 

Let $F$ be any field and suppose $X$ be a formal variable. 
Then $F[X]$ denotes the polynomial ring over $F$. 
We define the rational function field $F(X)$ to be the \emph{fraction ring} of $F[X]$ which consists of all quotients $\nicefrac{f}{g}$ where $f,g\in F[X]$ with $g(X)\neq 0$.
Similarly, we can define the polynomial ring $F[X_1,\dots, X_n]$ over multiple formal variables and their fraction field $F(X_1,\dots, X_n)$.

Look at the following example. $\bbR\subseteq \bbC$ are both fields. But $\bbC$ is a $2$-dimensional vector space over $\bbR$ as we can take the basis $\{1,i\}$.
So $\bbC/\bbR$ is a finite extension of degree $2$. 
But $\bbQ/\bbR$ is an infinite field extension. 
On the other hand let $\alpha\in \bbR$. Then consider the field $\bbQ(\alpha)$. 
Now degree of  $[\bbQ(\alpha)\colon \bbQ]$ depends on what kind of element $\alpha$ is. If $\alpha=\sqrt{2}$ then $[\bbQ(\sqrt{2})\colon \bbQ]=2$ but if $\alpha=\pi$ then $[\bbQ(\pi)\colon \bbQ]=\infty$.

\begin{Definition}{Generated Subfields}{def:generated-subfields}
  Let $K/F$ be a field extension and $S\subseteq K$.
  \begin{enumerate}[label=(\roman*)]
    \item The \emph{ring generated by $F$ and $S$}, written $F[S]$, is the intersection of all subrings of $K$ containing $F$ and $S$.
    \item The \emph{field generated by $F$ and $S$}, written $F(S)$, is the intersection of all subfields of $K$ containing $F$ and $S$.
  \end{enumerate}
  If $S=\{a_1,\ldots,a_n\}$ is finite we write $F[a_1,\ldots,a_n]$ and $F(a_1,\ldots,a_n)$. 
  We call $F(\alpha_1,\dots, \alpha_n)$ \emph{finitely generated} field. 
  For any field extension $K/F$ if there exists an element $\alpha\in K$ such that $K=F(\alpha)$ then such extension is called \emph{simple extension}.
\end{Definition}
From this definition it easily follows that for any $S\subseteq K$, $F(S)$ is the smallest subfield of $K$ containing $F$ and $\alpha$.

Let $K/F$ be a field extension. 
For any $\alpha\in K$ consider the map $\ev_{\alpha}:F[X]\to K$ defined by $f(X)\mapsto f(\alpha)$. 
Hence, $\ev_{\alpha}$ is a ring homomorphism.
\begin{observation}
  For any $\alpha\in K$, $\image(\ev_{\alpha})$ is a subring of $K$ and $\ev_{\alpha}$ is $F$-vector space homomorphism.
\end{observation}

Now we can use this map $\ev_{\alpha}$ to give descriptions of the elements of $F[\alpha]$ and $F(\alpha)$. 
\begin{lemma}{}{lem:simple-gen}
  Let $K$ be a field extension of $F$ and let $a\in K$. Then
  \[
    F[a] = \{f(a) : f(x)\in F[x]\}
    \quad\text{and}\quad
    F(a) = \left\{\frac{f(a)}{g(a)} : f,g\in F[x],\; g(a)\neq 0\right\}.
  \]
  Moreover, $F(a)$ is the quotient field of $F[a]$.
\end{lemma}

\begin{proof}
  The evaluation map $\mathrm{ev}_a$ is a ring homomorphism with image $\{f(a):f\in F[x]\}$/ 
  So this set is a subring of $K$. 
  Any subring of $K$ containing $F$ and $a$ must contain $f(a)$, for all $f\in F[X]$. 
  Therefore, $F[a]=\{f(a):f\in F[x]\}$. 
  The fraction field of $F[a]$ is $\{f(a)/g(a)\colon g(a)\neq 0\}$; it is contained in any subfield of $K$ containing $F[a]$, hence equals $F(a)$.
\end{proof}

We call this procedure \emph{adjoining $\alpha$} to $F$. 
Using similar arguments we can describe the ring $F[a_1,\dots, a_n]$ and the field $F(a_1,\dots, a_n)$ where $K/F$ is a field extension and $a_1,\dots, a_n\in K$. 
So we have the following theorem.
\begin{Theorem}{}{thm:mult-gen}
  Let $K$ be a field extension of $F$ and let $a_1,\ldots,a_n\in K$. Then
  \begin{align*}
    F[a_1,\ldots,a_n] &= \{f(a_1,\ldots,a_n) : f\in F[x_1,\ldots,x_n]\}, \\
    F(a_1,\ldots,a_n) &= \left\{\frac{f(a_1,\ldots,a_n)}{g(a_1,\ldots,a_n)} : f,g\in F[x_1,\ldots,x_n],\; g(a_1,\ldots,a_n)\neq 0\right\}.
  \end{align*}
  Moreover, $F(a_1,\ldots,a_n)$ is the quotient field of $F[a_1,\ldots,a_n]$.
\end{Theorem}

Now we can also define fields generated by adjoining arbitrary number of elements. 
In this case we can give the description of the new field in terms of all finite extension of fields created by adjoining finitely many elements. 

\begin{Theorem}{}{thm:fin-gen-sub}
  Let $K$ be a field extension of $F$ and let $S\subseteq K$ be any subset. 
  If $\alpha\in F(S)$, then $\alpha\in F(a_1,\ldots,a_n)$ for some $a_1,\ldots,a_n\in S$. 
  Therefore,
  \[
    F(S) = \bigcup\,\{F(a_1,\ldots,a_n) \colon a_1,\ldots,a_n\in S\},
  \]
  where the union is over all finite subsets of $S$.
\end{Theorem}

\begin{proof}
  Let $U\coloneqq \cup\,\{F(a_1,\ldots,a_n) \colon a_1,\ldots,a_n\in S\}$. 
  Each $F(a_1,\ldots,a_n)$ contains $F$ and lies in $F(S)$ where $a_1,\dots, a_n\in S$. 
  So the union $U$ satisfies $U\subseteq F(S)$. 
  The set $U$ contains $F$ and $S$. 
  If $\alpha\in F(a_1,\ldots,a_n)$ and $\beta\in F(b_1,\ldots,b_m)$, then $\alpha\pm\beta,\,\alpha\beta,\,\alpha/\beta$ (if $\beta\neq 0$) are all in $F(a_1,\ldots,a_n,b_1,\ldots,b_m)\subseteq U$. 
  Therefore, $U$ is a field. Being the smallest subfield containing $F$ and $S$, it equals $F(S)$.
\end{proof}

%------------------------------------------------------------------
\subsection{Algebraic Elements and Minimal Polynomials}
%------------------------------------------------------------------

\begin{Definition}{Algebraic and Transcendental Elements}{algebraic-elem}
  Let $K$ be a field extension of $F$. 
  An element $\alpha\in K$ is \emph{algebraic over $F$} if there is a nonzero polynomial $f(X)\in F[X]$ with $f(\alpha)=0$. 
  If $\alpha$ is not algebraic over $F$, then $\alpha$ is called \emph{transcendental over $F$}. 
  If every element of $K$ is algebraic over $F$, then $K$ is said to be \emph{algebraic over $F$}, and $K/F$ is called an \emph{algebraic extension}.
\end{Definition}

Hermite (1873) proved that $e$ is transcendental over $\mathbb{Q}$ and Lindemann (1882) proved the same for $\pi$. 
On the other hand, $\sqrt[n]{r}\in\mathbb{R}$ is algebraic over $\mathbb{Q}$ for any $r\in\mathbb{Q}$ (root of $x^n-r$), as is $i=\sqrt{-1}$ (root of $x^2+1$).

\begin{Definition}{Minimal Polynomial}{min-poly}
  If $\alpha$ is algebraic over a field $F$, the \emph{minimal polynomial} of $\alpha$ over $F$ is the monic polynomial $p(X)$ of least degree in $F[X]$ for which $p(\alpha)=0$; it is denoted $\min(F,\alpha)$. 
  Equivalently, $\min(F,\alpha)$ is the monic generator of the kernel of the evaluation homomorphism $\mathrm{ev}_\alpha$.
\end{Definition}

The minimal polynomial depends on the base field. 
For instance, $\min(\mathbb{Q},i)=x^2+1$ while $\min(\mathbb{C},i)=x-i$, since $i\in\mathbb{C}$ is already in the base field.

The minimal polynomial of an element and the degree of a field extension
are two of the most basic tools we shall use. 
The following proposition gives a relation between these objects.

\begin{Theorem}{}{thm:simple-ext}
  Let $K/F$ be a field extension and let $\alpha\in K$ be algebraic over $F$. 
  If $p(X)=\min(F,\alpha)$, then:
  \begin{enumerate}[label=\textup{(\roman*)}]
    \item $p(X)$ is irreducible over $F$.
    \item If $g(X)\in F[X]$, then $g(\alpha)=0$ if and only if $p(X)\mid g(X)$.
    \item If $n=\deg(\min(F,\alpha))$, then the elements $1,\alpha,\ldots,\alpha^{n-1}$ form a basis for $F(\alpha)$ over $F$. 
    So $[F(\alpha)\colon F]=\deg(\min(F,\alpha))<\infty$. Moreover, $F(\alpha)=F[\alpha]$.
    \item $F(\alpha)\cong \quotient{F[X]}{\la p(X)\ra}$.
  \end{enumerate}
\end{Theorem}
\begin{proof}
  The \emph{first isomorphism theorem} gives $F[\alpha]\cong \quotient{F[X]}{\la p(X)\ra}$. 
  Since $F[\alpha]\subseteq K$ is an integral domain, $\langle p(X)\rangle$ is a prime ideal. 
  Now in an \pid every nonzero prime is maximal, so $\quotient{F[X]}{\la p(X)\ra}$ is a field and $F[\alpha]=F(\alpha)$. 
  This gives (i) and (iv).  
  Part (ii) follows from $\ker(\ev_\alpha)=\langle p(X)\rangle$.

  For (iii): any $b\in F(\alpha)=F[\alpha]$ satisfies $b=g(\alpha)$. 
  Dividing gives $g=qp+r$ with $\deg r<n$ and $q,r\in F[X]$. 
  So $b=r(\alpha)$ lies in the span of $\{1,\alpha,\ldots,\alpha^{n-1}\}$. 
  Therefore, the set $\{1,\alpha,\dots, \alpha^{n-1}\}$ spans $F(\alpha)$ as a vector space. 
  If $a_0+a_1\cdot \alpha+\cdots +a_{n-1}\cdot \alpha^{n-1}=0$ where $a_0,\dots, a_{n-1}\in F$ then the polynomial $f(X)=\sum_{i=0}^{n-1}a_iX^i$ is in $\ker(\ev_\alpha)$. 
  Hence, either $f\equiv 0$ or $p\mid f$ by part (ii). 
  But $\deg(f)<n$. 
  So $f\equiv 0$. 
  Thus, $1,\alpha,\dots, \alpha^{n-1}$ forms a basis of $F(\alpha)$ over $F$. 
\end{proof}

The part (iii) of above theorem gives us a basis of $F(\alpha)$ over $F$ using only $\alpha$. 
So we get the following observation:
\begin{observation}
  If $K/F$ is a field extension and $\alpha\in K$ is algebraic over $F$ where $p=\min(F,\alpha)$ then $[F(\alpha)\colon F]=\deg(p)$. 
\end{observation}

Notice that if $\alpha,\beta\in K$ are both algebraic over $F$ and $\min(F,\alpha)=\min(F,\beta)=p$ i.e., their minimal polynomials are same then by the above theorem $$F(\alpha)\cong\quotient{F[X]}{\la p\ra}\cong F(\beta)$$
Therefore we have the observation:
\begin{observation}
  If $K/F$ is a field extension and $\alpha,\beta\in K$ are both algebraic over $F$, and they have the same minimal polynomials then $F(\alpha)\cong F(\beta)$ with the isomorphism $\varphi:\alpha\mapsto \beta$. 
\end{observation}
%------------------------------------------------------------------
\subsection{Finite and Algebraic Extensions}
%------------------------------------------------------------------

\begin{lemma}{}{thm:finite-alg}
  If $[K:F]<\infty$, then $K$ is algebraic and finitely generated over $F$.
\end{lemma}

\begin{proof}
  Let $\{\alpha_1,\ldots,\alpha_n\}$ be an $F$-basis of $K$. 
  Then for any element $\alpha\in K$, there exists $a_1,\dots, a_n$ such that $\alpha=a_1\cdot \alpha_1+\cdots +a_n\cdot \alpha_n$. 
  So certainly we have $K=F(\alpha_1,\dots, \alpha_n)$ and hence $K$ is finitely generated over $F$. 
  Since $[K\colon F]=n$, for any $\alpha\in K$, the $n+1$ elements $1,\alpha,\ldots,\alpha^n$ span only an $n$-dimensional space. 
  So they are $F$-linearly dependent. 
  Thus, there exists $\beta_0,\dots, \beta_n\in F$ such that $\sum_{i=0}^n\beta_i\cdot \alpha^i=0$. So we get a polynomial $f(X)=\sum_{i=0}^n \beta_i\cdot X^i$ and $f(\alpha)=0$. 
  Therefore, $\alpha$ is algebraic over $F$. 
\end{proof}

Let $K/F$ is a finite extension and $K=F(\alpha_1,\dots, \alpha_n)$.  
Then we can break up the extension $K/F$ into collection of subextensions. 
Let $L_i=F(\alpha_1,\dots,\alpha_i)$. 
Set $L_0=F$. 
Then basically $L_{i+1}=L_i(\alpha_{i+1})$. 
And we have $$F=L_0\subseteq L_1\subseteq \cdots \subseteq L_{n-1}\subseteq L_n=K$$
We will show a theorem which relates the degree of the extension $K/F$ and each subextensions. 

\begin{Theorem}{}{thm:tower-law}
  Let $F\subseteq L\subseteq K$ be fields. Then $[K:F]=[K:L]\cdot[L:F]$.
\end{Theorem}

\begin{proof}
  Let $\{a_i\}_{i\in I}$ be an $F$-basis of $L$ and $\{b_j\}_{j\in J}$ an $L$-basis of $K$. 
  We show $\{a_i\cdot b_j\}_{I\times J}$ is an $F$-basis of $K$. 
  For any $\gm\in K$, write $\gm=\sum_j\alpha_jb_j$ with $\alpha_j\in L$. 
  Now each $\alpha_j=\sum_i\beta_{ij}a_i$ with $\beta_{ij}\in F$; then $\gm=\sum_{i,j}\beta_{ij}a_ib_j$. 
  Therefore, the set $\{a_i\cdot b_j\}_{I\times J}$ spans $K$ over $F$. 
  
  If $\sum_{i,j}\beta_{ij}a_ib_j=0$ for some non-trivial linear combinations where $\beta_{i,j}\in F$ for all $i\in I$, $j\in J$. 
  Now $$\sum_{i,j}\beta_{ij}a_ib_j=\sum_j\bigl(\sum_i\beta_{ij}a_i\bigr)b_j=0.$$ 
  Since $\{b_j\}_{j\in J}$ is a basis of $K$ over $L$, each $\sum_i\beta_{ij}a_i=0$. 
  And by $F$-independence of $\{a_i\}$, each $\beta_{ij}=0$. 
  So $\{a_i\cdot b_j\}_{I\times J}$ form an $F$-basis of $K$. 
  Therefore, $$[K:F]=|\{a_i\cdot b_j\}_{I\times J}|=|\{a_i\colon i\in I\}|\cdot |\{b_j\colon j\in J\}|=[K\colon L]\cdot [L\colon F]$$ 
  So we have the theorem. 
\end{proof}

\begin{lemma}{}{thm:rat-func-deg}
  Let $K=F(X)$ be the field of rational functions in $X$ over $F$. 
  If $u\in K$ with $u\notin F$ let $u=\nicefrac{f}{g}$ where $f,g\in F[X]$ and $\gcd(f,g)=1$. 
  Let $L=F(u)$. 
  Then $$[K\colon L]=\max\{\deg(f),\deg(g)\}$$
\end{lemma}
\begin{proof}
  By \SimpleExtthm we need to find the minimal polynomial of $X$ over $L$ to determine $[K\colon L]$. 
  Consider the polynomial $p(Y)=u\cdot g(Y)-f(Y)\in L[Y]$. 
  Then $X$ is a root of $p$. 
  Therefore, $X$ is algebraic over $L$ and so $[K\colon L]<\infty$ and $K=L(X)$. 

  Notice that $\deg(p)=\max\{\deg(f),\deg(g)\}$. 
  Let $f=\sum_{i=0}^n a_iX^i$ and $g=\sum_{j=0}^m b_i\cdot X^i$. 
  We will show that $p$ is irreducible over $L$. 
  Now the element $u$ cannot be algebraic over $k$ otherwise by \TowerLawthm we have $[K\colon F]=[K\colon L]\cdot [L\colon F]<\infty$, and we know $F(X)$ is an infinite extension of $F$. 
  So $u$ is transcendental over $F$ and hence $F[u]\cong F[X]$. 
  Now observe that the polynomial $p\in F[Y][u]\subseteq F(Y)[u]$ which is of degree $1$. 
  Hence, $p$ is irreducible over $k(Y)$. 
  Now $p\in k[Y][u]=k[u][Y]$ and therefore $p$ is irreducible over $k(u)$. 
  Hence, $p$ is the minimal polynomial of $X$ over $F$. 
  So we have the lemma.
\end{proof}

\begin{note}
In the proof of \TowerLawthm the proof do not assume that the field extensions needs to be finite. 
\end{note}

We will now prove a converse theorem of \FiniteAlglem. 

\begin{Theorem}{}{thm:fin-gen-alg}
  Let $K/F$ be a field extension. 
  If $\alpha_1,\ldots,\alpha_n\in K$ are algebraic over $F$, then
  \[
    [F(\alpha_1,\ldots,\alpha_n)\colon F] \;\leq\; \prod_{i=1}^{n}[F(\alpha_i)\colon F].
  \]
\end{Theorem}

\begin{proof}
  We will prove this by induction on $n$. 
  The base case, $n=1$ follows from the \TowerLawthm. 
  So suppose this is true for $n-1$. 
  Take $L=F(\alpha_1,\ldots,\alpha_{n-1})$. 
  Then by induction hypothesis $[L:F]\leq\prod_{i=1}^{n-1}[F(\alpha_i):F]$. 
  Since $\alpha_n$ is algebraic over $F$ and hence over $L$, with $\min(L,\alpha_n)\mid\min(F,\alpha_n)$ in $L[X]$, we have $[L(\alpha_n):L]\leq[F(\alpha_n):F]$. 
  Therefore, $[F(\alpha_1,\dots, \alpha_n)\colon L]\leq [F(\alpha_n)\colon F]$. 
  So by \TowerLawthm we have 
  \[
    [F(\alpha_1,\ldots,\alpha_n):F] = [L(\alpha_n):L]\cdot[L:F] \leq [F(\alpha_n):F]\cdot\prod_{i=1}^{n-1}[F(\alpha_i):F].
  \]
  So we have the theorem. 
\end{proof}

The inequality above can be strict in some cases. 
For example let $a=\sqrt[4]{2}$ and $b=\sqrt[4]{18}$. 
Then both $[\bbQ(a)\colon \bbQ]=[\bbQ(b)\colon \bbQ]=4$ as $X^4-2$ and $X^4-18$ are both irreducible polynomials over $\bbQ$. 
But $[\bbQ(\sqrt[4]{2},\sqrt3)\colon \bbQ]=8$ and $\sqrt[4]{18}\in  \bbQ(\sqrt[4]{2},\sqrt3)$. 
Hence, in this the inequality doesn't follow. 

So from, and \FinGenAlgthm we have the following criterion for an element being algebraic over a field. 

\begin{corolary}{}{cor:alg-degree}
  Let $K/F$ be a field extension and $\alpha\in K$. 
  Then $\alpha$ is algebraic over $F$ if and only if $[F(\alpha):F]<\infty$. 
  Moreover, $K$ is algebraic over $F$ if $[K:F]<\infty$.
\end{corolary}

Now we can extend \FinGenAlgthm to the case of fields generated by arbitrary number of elements or extending field.
\begin{Theorem}{}{thm:alg-gen}
  Let $K$ be a field extension of $F$ and $X\subseteq K$ such that each element of $S$ is algebraic over $F$. 
  Then $F(S)/F$ is algebraic. 
  If $|S|<\infty$, then $[F(S):F]<\infty$.
\end{Theorem}

\begin{proof}
  For any $\alpha\in F(S)$. 
  \thmref{thm:fin-gen-sub} gives $\alpha\in F(a_1,\ldots,a_m)$ for some $a_1,\ldots,a_m\in S$. 
  By \thmref{thm:fin-gen-alg}, $F(a_1,\ldots,a_m)$ is algebraic over $F$. 
  So $\alpha$ is algebraic over $F$ by \thmref{thm:finite-alg}. 
  Since $\alpha$ is any arbitrary element of $F(S)$, $F(S)$ is algebraic over $F$. 
  If $S$ is finite, then $[F(S):F]<\infty$ by \FinGenAlgthm.
\end{proof}

We will now show that the property of being algebraic is a transitive property. In the finite extensions case this property easily follows from  \TowerLawthm and \AlgDegreecor. We will show this for general extensions below. 

\begin{Theorem}{Transitivity of Algebraic Extensions}{thm:alg-transitive}
  Let $F\subseteq L\subseteq K$ be fields. If $L/F$ and $K/L$ are algebraic, then $K/F$ is algebraic.
\end{Theorem}
\begin{proof}
  Let $\alpha\in K$. 
  Since $K/L$ is algebraic, let $f(X)=\sum_{i=0}^da_iX^i$ where each $a_i\in L$ is the minimal polynomial of $\alpha$ over $L$. 
  Since $L/F$ is algebraic and $f\in L[X]$, each $a_i$ is algebraic over $F$. 
  Consider the field $L'=F(a_0,\dots, a_d)$. 
  So by \thmref{thm:fin-gen-alg} the extension $L'/F$ is finite i.e.,  $[L'\colon F]<\infty$. 
  Since $f\in L'[X]$ and $p(\alpha)=0$, the element $\alpha$ is algebraic over $L'$ and $[L'(\alpha)\colon L']<\infty$. 
  By the tower law,
  \[
    [L'(\alpha)\colon F] = [L'(\alpha)\colon L']\cdot[L'\colon F] < \infty,
  \]
  Since $F(\alpha)\subseteq L'(\alpha)$, $\alpha$ is algebraic over $F$. 
  Since $\alpha$ is any arbitrary element of $K$, $K/F$ is algebraic. 
\end{proof}

\begin{Definition}{Algebraic Closure in $K$}{alg-closure-def}
  Let $K$ be a field extension of $F$. The set
  \[
    \{\,a\in K : a\text{ is algebraic over }F\,\}
  \]
  is called the \emph{algebraic closure of $F$ in $K$}.
\end{Definition}

\begin{corolary}{}{cor:alg-closure-field}
  Let $K$ be a field extension of $F$ and let $L$ be the algebraic closure of $F$ in $K$. Then $L$ is a field, and therefore is the largest algebraic extension of $F$ contained in $K$.
\end{corolary}

\begin{proof}
  Let $a,b\in L$.
  Then both are algebraic over $F$ by \thmref{thm:fin-gen-alg} and hence $F(a,b)$ is algebraic over $F$ and $F(a,b)\subseteq L$. 
  Since $a\pm b, a\cdot b, \nicefrac{a}{b}$ (if $b\neq 0$) all are in $F(a,b)\subseteq L$, $L$ is closed under the field operation. 
  So $L$ is a subfield of $K$. 
  Since each element of $K$ which is algebraic over $F$ is contained in $L$, $L$ is the largest algebraic extension of $F$ contained in $K$. 
\end{proof}

%------------------------------------------------------------------
\subsection{Splitting Field}
%------------------------------------------------------------------

\begin{Definition}{Splits over $K$}{def:splits}
  If $K$ is an extension field of $F$ and if $f(X)\in F[X]$, then $f$ \emph{splits over $K$} if $f(X)=a\prod_i(X-\alpha_i)\in K[X]$ for some $\alpha_1,\ldots,\alpha_n\in K$ and $a\in F$. In other words, $f$ splits over $K$ if $f$ factors completely into linear factors in $K[X]$.
\end{Definition}

In order to talk about roots of a given polynomial, we need to have
extension fields that contain the roots of the polynomial. 
The next theorem shows that for any $f\in F[X]$, there is a finite extension of $F$ over which $f$ splits. 

\begin{Theorem}{}{thm:ext-root-splitting}
  Let $f(x)\in F[x]$ have degree $n$. There is an extension field $K$ of $F$ with $[K:F]\leq n$ such that $K$ contains a root of $f$. In addition, there is a field $L$ containing $F$ with $[L:F]\leq n!$ such that $f$ splits over $L$.
\end{Theorem}
\begin{proof}
  Let $p(X)$ is an irreducible factor of $f(X)$ in $F[X]$ and let $K=\quotient{F[X]}{\la p(X)\ra}$. 
  Thus, $K$ contains a root of $p(X)$ say $\alpha$. 
  Therefore, $\alpha$ is also a root of $f(X)$. 
  Since $[K\colon F]=\deg(p)\leq n$ we have the first part. 

  For the second part we will use induction on $n$. 
  From the first part now we have a field $K\supseteq F$ such that $K$ contains a root $\alpha of F$. 
  So $f$ factors into $f(X)=(X-\alpha)g(X)$ in $K[X]$ where $g(X)\in K[X]$. 
  Now $\deg(g)=n-1$. 
  So by induction hypothesis there exists a finite extension $L$ of $K$ such that $[L\colon K]\leq (n-1)!$ and $g$ splits over $L$. 
  That means $f$ splits over $L$ and $[L\colon F]=[L\colon K]\cdot [K\colon F]\leq n!$. 
  So we have the theorem. 
\end{proof}

\begin{Definition}{Splitting Field}{}
  Let $K$ be an extension of $F$ and let $f\in F[X]$. 
  \begin{enumerate}[label=(\roman*)]
    \item $K$ is called the \emph{splitting field of $f$ over $F$} if and only if $f$ splits over $K$ and $K=F(\alpha_1,\dots, \alpha_n)$ where $\alpha_1,\dots, \alpha_n$ are exactly the roots of $f$. 
    \item If $S$ is a set of non-constant polynomials over $F$ then $K$ is the \emph{splitting field of $S$ over $F$} if and and only if  each $f\in S$ splits over $K$ and $K=F(S_R)$ where $S_R$ is the set of all roots of all $f\in S$. 
  \end{enumerate}
\end{Definition}

So \thmref{thm:ext-root-splitting} yields immediately the existence of splitting fields for a finite set of polynomials.
We have proved the existence of splitting fields for finite sets of polynomials. 
What about infinite sets? 
Suppose that $K$ is a splitting field over $F$ of the set of all non-constant polynomials over $F$. We do not know yet that such a field exists or not.

\begin{lemma}{}{lem:alg-closed-equiv}
  If $K$ is a field, then the following statements are equivalent:
  \begin{enumerate}[label={(\roman*)}]
    \item There are no algebraic extensions of $K$ other than $K$ itself.
    \item There are no finite extensions of $K$ other than $K$ itself.
    \item If $L$ is a field extension of $K$, then $K=\{a\in L : a\text{ is algebraic over }K\}$.
    \item Every $f(x)\in K[x]$ splits over $K$.
    \item Every $f(x)\in K[x]$ has a root in $K$.
    \item Every irreducible polynomial over $K$ has degree $1$. 
  \end{enumerate}
\end{lemma}
\begin{proof}We will show the implications in a circular order, (i)$\implies$(ii)$\implies$(iii)$\implies$(iv)$\implies$(v)$\implies$(vi)$\implies$(i). 
  \paragraph*{(i)$\implies$(ii):} This is clear since any finite extension is algebraic.
  \paragraph*{(ii)$\implies$(iii):} Let $a\in L$ is algebraic over $K$. 
  Then $K(a)$ is a finite extension of $K$. 
  But there are no finite extensions of $K$ other than itself. 
  So $K(a)=K$. 
  Thus, $a\in K$. 
  \paragraph*{(iii)$\implies$(iv):} Let $f\in K[X]$ and $L$ be the splitting field of $f$ over $K$. Since $L$ is algebraic over $K$ we have $L=K$ i.e., $f$ splits over $K$. 
  \paragraph*{(iv)$\implies$(v):} This is obviously true. 
  \paragraph*{(v)$\implies$(vi):} Let $f\in K[X]$ is irreducible. 
  Therefore, $f$ has a root in $K$, so $f$ has a linear factor. 
  But $f$ is irreducible. 
  Hence, $\deg(f)=1$. 
  \paragraph*{(vi)$\implies$(i):} Let $L$ be an algebraic extension of $K$.
  Let $a\in L$ and $p=\min(K,a)$. 
  Then $\deg(p)=1$. 
  So $[K(a)\colon K]=1$ and hence $a\in K$. 
  So $L=K$. 
\end{proof}

\begin{Definition}{Algebraically Closed Field}{}
  If any field $K$ satisfies the equivalent conditions of \lemref{lem:alg-closed-equiv} then $K$ is said to be \emph{algebraically closed}. If $K$ is an algebraic extension of $F$ then $K$ is said to be \emph{algebraic closure} of $F$. 
\end{Definition}

We haven't showed the existence of such theorem or even every field as an algebraic closure. Consider the field $\bbC$. This field is algebraically closed. But we will not prove if every field has an algebraic closure. We will just state the theorem and move on. 
\begin{Theorem}{}{}
  Let $F$ is a field. Then $F$ has an algebraic closure. 
\end{Theorem}

The existence of an algebraic closure yields immediately the existence of a splitting field for an arbitrary set of non-constant polynomials.
\begin{corolary}{}{}
  Let $F$ be a field and $S$ be a set of non-constant polynomials over $F$. 
  Then $S$ has a splitting field over $F$. 
\end{corolary}
\begin{proof}
  Let $K$ be an algebraic closure of $F$. Then each $f(x)\in S$ splits over $K$. Let $X$ be the set of roots of all $f\in S$. Then $F(X)\subseteq K$ is a splitting field for $S$ over $F$, since each $f$ splits over $F(X)$ and this field is generated by the roots of all the polynomials from $S$.
\end{proof}

So using the corollary we can immediately relate algebraic closure with splitting fields. 
\begin{corolary}{}{}
  If $F$ is a field then the splitting field of the set of all non-constant polynomials over $F$ is an algebraic closure of $F$. 
\end{corolary}

But now comes another question. Can one field have multiple algebraic closures? Or even can we have more than one splitting fields? 
The answer is splitting fields are unique, and hence algebraic closures are also unique. 
Below we share the theorem which proves the above. We will not prove it because its beyond our scope of study. But for further reading and understanding the reader can look \cite{Morandi1996}.

\begin{Theorem}{Isomorphism Extension Theorem, \cite[Chapter 1]{Morandi1996}}{thm:iso-ext}
  Let $\sigma:F\to F'$ be a field isomorphism. Let $S=\{f_i(x)\}$ be a set of polynomials over $F$, and let $S'=\{\sigma(f_i)\}$ be the corresponding set over $F'$. Let $K$ be a splitting field for $S$ over $F$, and let $K'$ be a splitting field for $S'$ over $F'$. Then there is an isomorphism $\tau:K\to K'$ with $\tau|_F=\sigma$. Furthermore, if $\alpha\in K$ and $\alpha'$ is any root of $\sigma(\min(F,\alpha))$ in $K'$, then $\tau$ can be chosen so that $\tau(\alpha)=\alpha'$.
\end{Theorem}

The isomorphism extension theorem immediately settles the question of uniqueness of splitting fields and algebraic closures: any two splitting fields of the same set of polynomials over $F$ must be isomorphic, since the identity on $F$ extends to an isomorphism between them. As a special case, algebraic closures are unique up to $F$-isomorphism.

\begin{corolary}{}{cor:splitting-unique}
  Let $F$ be a field, and let $S$ be a subset of $F[x]$. Any two splitting fields of $S$ over $F$ are $F$-isomorphic. In particular, any two algebraic closures of $F$ are $F$-isomorphic.
\end{corolary}

\begin{proof}
  For the proof of the first statement, the isomorphism extension theorem gives an isomorphism extending $\mathrm{id}$ on $F$ between any two splitting fields of $S$. The second statement follows from the first, since any algebraic closure of $F$ is a splitting field of the set of all non-constant polynomials in $F[x]$.
\end{proof}

As a corollary to the existence and uniqueness of algebraic closures, we can prove that any algebraic extension of a field $F$ can be viewed as living inside a fixed algebraic closure of $F$.

\begin{corolary}{}{cor:alg-ext-in-closure}
  Let $F$ be a field, and let $N$ be an algebraic closure of $F$. If $K$ is an algebraic extension of $F$, then $K$ is isomorphic to a subfield of $N$.
\end{corolary}

\begin{proof}
  Let $M$ be an algebraic closure of $K$. By \thmref{thm:alg-transitive}, $M$ is algebraic over $F$; hence, $M$ is also an algebraic closure of $F$. Therefore, by the previous corollary, $M\cong N$. If $f:M\to N$ is an $F$-isomorphism, then $f(K)$ is a subfield of $N$ isomorphic to $K$.
\end{proof}