\section{Some Properties of Finite Fields}
In this section we describe fields of finite order. 
The characteristic of a finite field cannot be zero so it is a prime number. 
Therefore, if $K$ is a finite field and $\Char(K)=p$ then $\bbF_p\subseteq K$. 
Since $K$ is finite it will be a finite dimensional vector space over $\bbF_p$. 
So $K$ is isomorphic to $\bbF_p^r$ where $r=[K\colon \bbF_p]$ as a $\bbF_p$-vector space which contains $p^r$ elements. 
So order of a finite field is always a prime power. 
We write $\bbF_q\coloneqq K$ where $q=p^r$.
Below we list the main facts of finite fields without proof. 
If the reader is interested he can look into the book \cite{Artin2011}.

\begin{Theorem}{Properties of Finite Field}{thm:finite-field-props}
  Let $p$ be a prime integer, and let $q=p^r$ be a positive power of $p$.
  \begin{enumerate}[label=(\roman*)]
    \item Let $K$ be a field of order $q$. The elements of $K$ are roots of the polynomial $X^q-X$.
    \item The irreducible factors of the polynomial $X^{q}-X$ over the field $\bbF_p$ are the irreducible polynomials in $\bbF_p[X]$ whose degrees divide $r$.
    \item Let $K$ be a field of order $q$. The multiplicative group $K^*$ of nonzero elements of $K$ is a cyclic group of order $q-1$.
    \item There exists a field of order $q$, and all fields of order $q$ are isomorphic.
    \item A field of order $p^r$ contains a subfield of order $p^k$ if and only if $k$ divides $r$.
  \end{enumerate}
\end{Theorem}

\begin{observation}\label{obs:finite-ext-simple}
  Since for every $r\in\bbN$ there exists an element $\alpha\in \bbF_{q^r}$ such that $\bbF_{q^r}^*$ is generated by $\alpha$ then $\bbF_{q^r}=\bbF_q(\alpha)$. Hence, every finite extensions of finite fields are simple extensions. 
\end{observation}

Since finite fields are finite extensions of the base prime field $\bbF_p$ we can obtain irreducible polynomials for every positive integer degree. 

\begin{corolary}{}{cor:every-degree-irr-fp}
  For every $d\in\bbN$ there exists an irreducible polynomial of degree $d$ over the finite field $\bbF_p$.
\end{corolary}
\begin{proof}
  By the part (iv) of the above theorem here is a field $K$ of order $p^d$. Its degree $[K\colon \bbF_p]=d$. By part (iii) there is an element $\alpha\in K$ such that $\alpha$ generates the cyclic group $K^*$. So $K=\bbF_p(\alpha)$. Therefore, $\alpha$ is a root of a $d$-degree irreducible polynomial over $\bbF_p$. 
\end{proof}

\begin{Definition}{Primitive Element}{}
  Let $K$ be finite field over $\bbF_p$. Then $K$ has an element $\alpha$ which generates $K/F$ i.e., $K=\bbF_p(\alpha)$. Then $\alpha$ is called the \emph{primitive element} of $K$. 
\end{Definition}

Let $f$ is an irreducible polynomial over the finite field $\bbF_q$ and let $\alpha$ be a root of $f$ in the splitting field of $f$ over $\bbF_q$. 
Therefore, $f=\min(\bbF_q,\alpha)$. Now suppose $h\in\bbF_q[X]$ such that $h(\alpha)=0$. So by \SimpleExtthm, $f\mid h$. Like \thmref[(ii)]{thm:finite-field-props} we will show that we can similar result for any general finite field. 
\begin{lemma}{}{lem:irr-divides-frobenius}
  Let $f\in\bbF_q[X]$ be an irreducible polynomial of degree $d$.
  Then $f(X)$ divides $X^{q^n}-X$ if and only if $d\mid n$. 
\end{lemma}
\begin{proof}
  Suppose $f\mid X^{q^n}-X$. Let $\alpha$ be a root of $f$ in the splitting field. 
  Then $\alpha^{q^n}=\alpha$ and so $\alpha\in\bbF_{q^n}$. 
  Now $\bbF_q\subseteq \bbF_q(\alpha)\subseteq \bbF_{q^n}$. 
  Since $[\bbF_q(\alpha)\colon \bbF_q]=d$ and $[\bbF_{q^n}\colon \bbF_q]=n$. So by \TowerLawthm, $d\mid n$.

  Let $d\mid n$. Then $\bbF_{q^d}\subseteq \bbF_{q^n}$. If $\alpha$ is a root of $f$ in the splitting field of $f$ over $\bbF_q$ then $\bbF_q(\alpha)=\bbF_{q^d}\subseteq \bbF_{q^n}$. Therefore, $\alpha$ is a root of $X^{q^{n}}-X$. Since $f$ is the minimal polynomial of $\alpha$ and $X^{q^n}-X\in\bbF_q$, $f\mid X^{q^n}-X$. 
\end{proof}

\begin{Theorem}{}{thm:irr-roots-conjugates}
  If $f$ is an irreducible polynomial of $\bbF_q[X]$ of degree $d$ then $f$ has a root $\alpha\in\bbF_{q^d}$. Furthermore, all roots of $f$ are given by $d$ distinct elements $\alpha,\alpha^q,\alpha^{q^2},\dots, \alpha^{q^{d-1}}$ of $\bbF_{q^d}$.
\end{Theorem}
\begin{proof}
  Let $f(X)=\sum_{i=0}^da_iX^i$. 
  Then $$f\lt(\alpha^{q^j}\rt)=\sum_{i=0}^da_i\cdot \lt(\alpha^{q^j}\rt)^i=\sum_{i=0}^da_i^{q^j}\cdot \lt(\alpha^i\rt)^{q^j}=\lt(\sum_{i=0}^d a_i\cdot \alpha^i\rt)^{q^j}=f^{q^j}(\alpha)=0$$Since $\alpha^{q^{d}}=\alpha$ the roots of $f$ are $\alpha,\alpha^q,\dots, \alpha^{q^{d-1}}$. 
\end{proof}

This introduces us to look at roots and field automorphisms in a new way. Let $F,F'$ are two fields and $\vph:F\to F'$ is a homomorphism.
Then there is an induced homomorphism $\vph:F[X]\to F'[X]$ given by $$\vph\lt(\sum_{i=0}^na_i\cdot X^i\rt)=\sum_{i=0}^n \vph(a_i)\cdot X^i$$
So if $f=\sum_{i=0}^na_iX^i$ factors into $a(X-\alpha_1)\cdots (X-\alpha_n)$ then $\vph(f)=\vph(a)(X-\vph(\\alpha_1))\cdots (X-\vph(\alpha_n))$. The \emph{Frobenius map},  $\sg:\alpha\mapsto \alpha^q$ for any element of the extended field $\bbF_{q^d}$ over $\bbF_q$ is a $\bbF_q$-automorphism meaning that automorphism fixes the elements of $\bbF_q$. The above theorem tells us if $\alpha$ is a root of $f$ which is an irreducible polynomial over $\bbF_q$ then $\sg(\alpha)$ is also a root.

Now by the above theorem if $f$ is irreducible over $\bbF_q$ of degree $d$ then $\bbF_{q^d}$ contains a root $\alpha$ of $f$. Then $\alpha,\alpha^q,\dots, \alpha^{q^{d-1}}$ all are in $\bbF_{q^d}$. Therefore, $f$ splits in $\bbF_{q^d}$. Since $f$ is the minimal polynomial of $\alpha$, $\bbF_{q^d}=\quotient{\bbF_q[X]}{\la f\ra}$. So $\bbF_{q^d}$ is the splitting field of $f$. Since all finite fields of size $q^d$ are isomorphic for any other degree $d$ irreducible polynomial $g\in\bbF_q[X]$, if $\beta$ is a root of $g$ then $$\bbF_q(\alpha)\cong \quotient{\bbF_q[X]}{\la f\ra}\cong\bbF_{q^d}\cong \quotient{\bbF_q[X]}{\la g\ra}\cong\bbF_q(\beta)$$So the splitting fields of $f$ and $g$ are isomorphic. So we have the following corollary.

\begin{corolary}{}{cor:irr-splitting-field-fqd}
  Let $f$ is an irreducible polynomial of $\bbF_q[X]$ of degree $d$. Then the splitting field of $f$ over $\bbF_q$ is given by $\bbF_{q^d}$. Moreover, any two irreducible polynomials $f,g\in\bbF_q[X]$ of same degree has isomorphic splitting fields.
\end{corolary}

Now we have heard the term conjugate of an element in complex numbers. The notion of conjugates is more general than just in complex numbers. 
\begin{Definition}{}{}
  Let $\bbF_{q^d}$ be an extension of $\bbF_q$ and $\alpha\in\bbF_{q^d}$. Then the elements $\alpha,\alpha^q,\dots, \alpha^{q^{d-1}}$ are called the \emph{conjugates} of $\alpha$. 
\end{Definition}

The conjugates of $\alpha\in\bbF_{q^m}$ with respect to $\bbF_q$ are distinct if and only if the minimal polynomial of $\alpha$ over $\bbF_q$ has degree $m$. Otherwise, the degree $d$ of this minimal polynomial is a proper divisor of $m$, and then the conjugates of $\alpha$ with respect to $\bbF_q$ are the distinct elements $\alpha,\alpha^q ,\dots,\alpha^{q^{d-1}}$, each repeated $\nicefrac{m}d$ times.
\begin{lemma}{}{lem:conjugates-same-order}
  For any $\alpha\in\bbF_{q}^*$ the conjugates with respect to any subfield of $\bbF_q$ have the same order in the group $\bbF_q^*$.
\end{lemma}
\begin{proof}
  Let $\bbF_{q'}$ be a subfield of $\bbF_q$ where $q'^r=q$. Since $\bbF_q^*$ is a cyclic group of order $q-1$ and $\gcd(q'^i,q-1) =1$ for all $0\leq i\leq r-1$. Therefore, the order of the conjugates of $\alpha$ has the same order as $\alpha$ in the group $\bbF_q^*$.  
\end{proof}