\chapter*{Acknowledgements}
I would like to thank my guide Prof. Mrinal Kumar for guiding me through out the project. Over the course of numerous meetings, he has cleared concepts, doubts, and difficulties that I encountered, not only in this project but also during different stages of my academic life for which he has my warmest thanks. 

I would also like to sincerely thank Prof. Prahladh Harsha for his guidance and support throughout this period. The discussions and meetings with Prof. Prahladh Harsha and Prof. Mrinal Kumar have greatly shaped my understanding and research. He consistently suggested interesting topics and papers to read, encouraged me to learn new areas, and guided me whenever I struggled with unfamiliar concepts. Beyond academics, their advice, and encouragement during moments of self-doubt have helped me immensely. I am deeply grateful for his kindness, patience, and support in many aspects throughout this journey.

I would also like to thank Prof.\ Ramprasad Saptharishi, Prof.\ Arkadev Chattopadhyay, Prof.\ Umang Bhaskar, and Prof.\ T. Kavitha for their courses at TIFR, which introduced me to several beautiful ideas in mathematics and theoretical computer science. Equally, I am grateful to Prof.\ Partha Mukhopadhyay, Prof.\ Amit Kumar Sinhababu, and Prof.\ Sumanta Ghosh for the foundations they built during my time at CMI --- their teaching shaped my love for the subject more than they perhaps know.

I would like to express my deepest gratitude to my parents for their unconditional love, sacrifices, and constant support throughout my academic journey. Despite the many hardships and medical difficulties faced by both my mother and father, they always encouraged and supported my love for mathematics and theoretical computer science. I especially want to acknowledge my mother’s strength and resilience in enduring serious medical hardships while continuing to care for and encourage me at every stage. Their unwavering belief in me, even during difficult times, has been a constant source of strength, and whatever I have achieved would not have been possible without them.

Nishant Das, Shubham Bhardwaj, Spandan Podder, and Vivek Karunakaran have made my time at TIFR genuinely special. From arguments about computer science that drifted into everything else, to the kind of easy honesty that only comes with real friendship, they have been a constant presence --- the sort of people whose company makes a difficult period feel manageable and, often, quite good. I did not expect to find friends like these here, and I am glad I did.

I also want to thank the postdoctoral researchers at TIFR --- Susobhon, Sagnik, Dipan, and Anirban --- for their warmth and for advice about research life that was always as honest as it was welcome.

Saptarshi Sahoo, Sougata Panda, and Soumil Baksi are among the closest people in my life. What we had at CMI --- the hours we lost track of, the arguments that had no end, the shared chaos of the S24 Room --- is something I carry with me, and they remain my dearest friends to this day. I am grateful for them in a way that goes well beyond what these pages can hold.

The CMI Zone deserves a mention of its own --- that little corner where we spent countless hours, and where so many of the best moments quietly happened.

Finally, I want to specially thank Somnath Bhattacharjee and Saswata Mukherjee, whose enthusiasm for the algebraic side of theoretical computer science was infectious and did a great deal to shape where my interests eventually led. 