\chapter{Exponential Sums over Finite Fields}
Exponential sums are important tools in number thoery for solving problems involving integers. These sums have many applications in Analytic Number Theory. 

In this chapter we will set up exponential sums for finite fields by special group homomorphisms called \textit{characters}. Since a finite field has two different inherent group structure; one is additive and the other is multiplicative we obtain two different characters. Exponential sums are formed by using the values of one or more characters and possibly combining them with weights or with other function values.
\section{Characters of Finite Abelian Groups}
Let $G$ be a finite abelian group with the identity element $1_G$. Let $S^1$ be the set of complex numbers with absolute value $1$. Then consider the multiplicative group $S^1$ with the operation of multiplication. 
\begin{Definition}{Character $\chi$ of $G$}{}
  A \textit{character} $\chi$ of $G$ is a group homomorphism $\chi:G\to S^1$ to the multiplicative group $S^1$ i.e. for all $g_1,g_2\in G$ $$\chi(g_1\cdot g_2)=\chi(g_1)\cdot \chi(g_2)$$
\end{Definition}

So if $\chi$ is any character of $G$ then $\chi(1_G)=1$. Since by definition for all $g\in G$, $$\chi(g)^{|G|}=\chi(g^{|G|})=\chi(1_G)=1$$ the values of $\chi$ are $|G|^{th}$ roots of unity. Also we have for any $g\in G$, $$\chi(g)\chi(g^{-1})=\chi(g\cdot g^{-1})=\chi(1_G)=1$$ so $\chi(g^{-1})=\chi(g)^{-1}=\overline{\chi(g)}$.

There are some special characters of $G$ which have specific name. The character $\chi_0$ defined by $\chi_0(g)=1$ for all $g\in G$ is called the \textit{trivial character}. For any character $\chi$ of $G$, the character $\overline{\chi}$ defined by $\overline{\chi}(g)=\overline{\chi(g)}$ for all $g\in G$ is called the \textit{conjugate character} of $\chi$. 

Given any finitely many characters of $G$, $\chi_1,\dots, \chi_n$ we can define the product character $\chi=\chi_1\cdot \chi_2\cdots \chi_n$ such that for all $g\in G$, $$\chi(g)=\chi_1(g)\cdot \chi_2(g)\cdots \chi_n(g)$$
Hence for all $g\in G$, $$\chi(g)\cdot \overline{\chi}(g)=|\chi(g)|^2=1$$Hence the $\overline{\chi}$ is also the inverse character of $\chi$. Hence the set of all characters of $G$ forms a group under multiplication whose identity element is the trivial character $\chi_0$. 
\begin{Definition}{Dual Group to $G$}{}
  For a finite abelian group $G$, the group of all characters of $G$ with the identity element $\chi_0$ is called the \textit{dual group} to $G$ and is denoted by $\widehat{G}$.
\end{Definition}

\begin{observation}
The dual group $\widehat{G}$ is abelian. 
\end{observation}
Since the values of characters of $G$ can only be $|G|^{th}$ roots of unity, the dual group $\widehat{G}$ is a finite abelian group. 
\begin{lemma}{}{lem:chars-cyclic-group}
  If $G$ is a finite cyclic group of order $n$ and let $g$ be the generator of $G$, then for all $j\in\{0,1,\dots, n-1\}$ define the function $$\chi_j(g^k)=e(\nicefrac{jk}{n}) \quad \forall\ k\in \{0,1,\dots, n-1\}$$Then $\hG=\{\chi_j\colon j\in\{0,1,\dots, n-1\}$. 
\end{lemma}
\begin{proof}
  Certainly for all $j\in\{0,1,\dots, n-1\}$ defines a character of $G$. On the other hand suppose, $\chi\in\hG$. Then $\chi(g)$ must be an $n^{th}$ root of unity. Hence there exists $j\in\{0,1,\dots, n-1\}$ such that $\chi(g)=e(\nicefrac{j}{n})$. Therefore $\chi=\chi_j$. Hence $\hG=\{\chi_j\colon  0\leq j\leq n-1\}$. 
\end{proof}
\begin{observation}
  The characters defined in above lemma forms a cyclic group with $\chi_1$ being the generator of $\hG$. 
\end{observation}
\begin{corolary}{}{}
  If $G$ is a finite cyclic group of order $n$ with $g$ being the generator then $G\cong \hG$ by the isomorphism $\varphi:g^j\mapsto \chi_j$ for all $j\in\{0,1,\dots, n-1\}$. 
\end{corolary}

Now we will show that we can extend a character of any subgroup to a character of the group. 
\begin{Theorem}{}{thm:subgroup-char-extend}
  Let $H$ be a subgroup of the finite abelian group $G$ and let $\chi_H$ be a character of $H$. Then $\chi_H$ can be extended to a character of $G$ that is there exists a character $\chi\in\hG$ with $\chi(h)=\chi_H(h)$ for all $h\in H$. 
\end{Theorem}
\begin{proof}
  Let $H$ be a proper subgroup of $G$. Let $\alpha\in G\setminus H$. Then let $H_1$ be the group generated by $H$ and $\alpha$ i.e. $H_1=\langle H,a\rangle$. Let $m$ be the smallest positive integer such that $\alpha^m\in H$. Then for all $g\in H_1$, $g$ can be written uniquely as $g=\alpha^j\cdot h$ where $0\leq j<m$ and $h\in H$. 

  Let $\om$ be a complex number such that $\om^m=\chi_H(\alpha^m)$. Then for all $g\in H_1$ where $g=\alpha^j\cdot h$, $h\in H$ define the function $\chi_1(g)=\om^j \chi_H(h)$. Then for any $g_1,g_2\in H_1$ where $g_1=\alpha^i\cdot h_1$, $g_2=\alpha^j\cdot h_2$, we have $g_1\cdot g_2=\alpha^{j+k}\cdot h_1\cdot h_2=\alpha^{j+k\bmod{m}}\cdot h_1\cdot h_2$. Hence $$\chi_1(g_1\cdot \chi_2)=\om^{j+k \bmod{m}}\chi_H(h_1)\cdot \chi_H(h_2)=\alpha^j\chi_H(h_1)\cdot \om^k\chi_H(h_2)=\chi_1(g_1)\cdot \chi_1(g_2)$$Hence $\chi_1$ is a character of $H_1$. And also for all $h\in H$ we have $\chi_1(h)=\chi_H(h)$. 
  
  So if $H_1=G$ we are done. Otherwise we can continue this process above until we get $G$. This process we have to do finitely many times as $G$ is a finite group.
\end{proof}


Therefore if we have two distinct elements $g_1,g_2\in G$, $g_1\neq g_2$ then take the subgroup generated by the element $h=g_1\cdot g_2^{-1}\neq 1$. Since this subgroup is a cyclic subgroup generated by $h$, by \lmref{lem:chars-cyclic-group} we get a character $\chi'$ of the cyclic subgroup such that $\chi'(h)\neq 1$. Then by \thmref{thm:subgroup-char-extend} we get a character $\chi$ of $G$ such that $\chi(h)=\chi'(h)\neq 1$ and hence $\chi(g_1)\neq \chi(g_2)$. Therefore we have the following observation
\begin{observation}
For any two distinct elements $g_1,g_2\in G$, $g_1\neq g_2$ we can get a character $\chi\in\hG$ such that $\chi(g_1)\neq \chi(g_2)$. 
\end{observation} 

Now every element of $\hG$ is a function $\chi$ which takes an element of $G$ and maps to a complex number in $S^1$. Similarly we can think of every element of as a function from $\hG$ to $S^1$. In fact for every element $g\in G$ define the homomorphism $\ev_g:\hG\to S^1$ where $\varphi_g(\chi)=\chi(g)$. 
\section{Gauss Sums}
\section{Jacobi Sums}
\section{\texorpdfstring{$L$}{L}-functions}