\chapter{Decoding from Character Evaluations}

Character sums of polynomials over finite fields carry deep arithmetic information, and one of the most striking manifestations of this is that they give rise to good error-correcting codes. The mechanism is simple: two polynomials whose character evaluations are close must be nearly indistinguishable by the character, and the Weil bounds make this precise. For a multiplicative character $\psi$ of $\bbF_q$ and two polynomials $f, g \in \bbF_q[X]$ of small degree, the Weil bound controls gives a lower bound on the Hamming distance between the codewords $(\psi(f(\alpha)))_{\alpha \in \bbF_q}$ and $(\psi(g(\alpha)))_{\alpha \in \bbF_q}$. The resulting codes turn out to have distance approaching $q/2$ while the message length grows polynomially in $q$.

Two families are of particular interest. For $q$ odd, applying the quadratic residue character $\psi(\beta) = \beta^{\nicefrac{(q-1)}2}$ to squarefree monic polynomials $g(X)$ of degree $d = o(\sqrt{q})$ defines a nonlinear code over $\{0,\pm 1\}$ of length $q$. For $q = 2^b$, taking $g(X)$ with only odd-degree monomials and forming $\tr \circ g : \bbF_q \to \bbF_2$ gives the classical \emph{dual BCH codes}, introduced independently by Bose--Ray-Chaudhuri \cite{Bose1960} and Hocquenghem \cite{Hoc59}. These codes are linear over $\bbF_2$, have length $q$ and dimension $\frac{d+1}{2}\log_2 q$, and their distance analysis relies entirely on the Weil bound for additive character sums. Dual BCH codes are remarkable: for small $d$ they are \emph{unmatched} codes in the sense that, without the Weil bounds, there is no known proof that binary codes of their parameters even exist.

Despite this theoretical richness, decoding these codes from a constant fraction of errors was a long-standing open problem. The essential difficulty is that each evaluation $\psi(g(\alpha))$ carries very little information about $g$ --- the quadratic character discards the magnitude and retains only whether $g(\alpha)$ is a quadratic residue --- so the system of equations one must solve is both underdetermined and nonlinear. Even the zero-error interpolation problem, of recovering $g$ exactly from the values $\psi(g(\alpha))$ for all $\alpha \in \bbF_q$, had no polynomial-time solution for non-constant $d$ prior to the work presented here.

The goal of this chapter is to present the first polynomial-time algorithms for decoding both code families from a constant fraction of errors, following Kopparty. The algorithms combine two classical ideas: the \emph{Berlekamp--Welch} framework, which uses a high-multiplicity error-locator polynomial to zero out error positions and reduces decoding to a linear system, and the \emph{Stepanov polynomial method}, the elementary approach to the Weil bounds that works by interpolating auxiliary polynomials vanishing with high multiplicity at points of interest. The fusion of these two ideas gives rise to what we call \emph{pseudopolynomials}: high-degree polynomials all of whose Hasse derivatives, when evaluated on $\bbF_q$, agree with polynomials of much smaller degree. These objects turn out to be the correct algebraic framework for capturing the special structure of character evaluations of polynomials.

\input{chapters/decoding/pseudopolynomials}
\input{chapters/decoding/quadratic-character-eval}