\section{Decoding Multiplicative Character Evaluations for Odd Characteristic}
Let $\psi\in \Xq$ is a non-trivial multiplicative character of $\bbF_q$ where $\ord(\psi)=d$. Then the polynomial based code we can construct is $\lt(\psi\circ f(\alpha)\rt)_{\alpha\in \bbF_q}$ where $f\in\bbF_q[X]$. Now suppose $g$ is another polynomial. Then for any $\alpha\in\bbF_q$, $$\psi((f\cdot g^d)(\alpha))=\psi(f(\alpha))\cdot \psi^d(g(\alpha))=\psi(f(\alpha))$$So we assume $f$ doesn't have a power $d$ factor. Similarly, for any $\alpha,\beta$ where $\beta\neq 0$ $$\psi((\beta^{d}\cdot f)(\alpha))=\psi^d(\beta)\cdot \psi(f(\alpha))=\psi(f(\alpha))$$Since different leading coefficient can lead to same character evaluation we assume $f$ to be monic.

\subsection{Quadratic Character}
Now we take the quadratic character $\eta\in\Mq$. Hence, we assume the polynomials are monic and squarefree. Then by \WeilBoundPsiNondthm we have $$\lt|\sum_{\alpha}\eta((g_1\cdot g_2)(\alpha))\rt|< O(d\sqrt{q})$$Therefore the vectors $\eta\circ g_1$ and $\eta\circ g_2$ differs in at least $\lt(q-O(d\sqrt{q})\rt)=\lt(\frac12-\frac{d}{\sqrt{q}}\rt)q$ many locations. So their hamming distance is at least $\lt(\frac12-\frac{d}{\sqrt{q}}\rt)q$ which gives the distance of the code.

Below we give an algorithm to decode these codes from $\nicefrac14$th of the minimum distance of the code. The decoding algorithm follows the \emph{Berlekamp-Welch} of \emph{Reed-Solomon codes} to some extent. Let $g$ is the closest codeword. So suppose $G(X)\coloneqq g^{\nicefrac{(q-1)}2}(X)$.

If there had been zero error in the received word then we know the polynomial $\tdG$ is $\psi\circ g$-twisted $(O(dM),M))$-pseudopolynomial. But we have the received word very close to the original codeword $\psi\circ g$. So we will try to find a $r$-twisted pseudopolynomial $F$ with good parameters such that we can use \TwistedTwolem or \TwistedRobustTwothm to relate them. So first we have to show such $F$ exists.

\subsubsection{Constructing \texorpdfstring{$r$}{r}-twisted Pseudopolynomial}
Let $S=\{\alpha\in\bbF_q\colon \psi\circ g(\alpha)\neq r(\alpha)\}$ be the error locator set. Then consider the error locator polynomial $Z_S(X)=\prod_{\alpha\in S}(X-\alpha)^M$. Then all elements of $S$ are roots of $Z_S$ with multiplicity at $M$. So we take any multiple of $Z_S$, $E(X)$ which is a $h$-pseudopolynomial of degree $<c$ where $h,c,M$ are parameters to be chosen later. So by construction $E(X)=\sum_{i=0}^{c-1}E_i(X)\cdot \Lambda^i(X)$ by \PdegCharlem. So $\deg(E_i)\leq h$. Therefore the number of coefficients is $c(h+1)$. But number of constraints is $e\cdot M$. Hence we want $c(h+1)>e\cdot M$. So it is enough to have $ch=eM$.

So now we have a nonzero multiple of $Z_S(X)$ such that for all $\alpha\in S$, $\mult(E,\alpha)\geq M$. Therefore we take the polynomial $G(X)\cdot E(X)$ where $G(X)=g^{\frac{q-1}m+M}(X)$. We will show that this $F$ is a $r$-twisted $(h+dM,M)$-pseudopolynomial.

First we will calculate $E_{\la \ell\ra}$ for all $0\leq \ell<M$. By \DerivativeXqXlem we have $$E^{(\ell)}(X)=\sum_{i=0}^{c-1}(-1)^iE_i^{(\ell -i)}(X)\bmod{\Lm(X)}\implies E_{\la \ell\ra}(X)=\sum_{i=0}^{c-1}(-1)^iE_i^{(\ell -i)}(X)$$Therefore, for any $\alpha\in\bbF_q$ we have \begin{align*}
  (E\cdot G)^{(\ell)}(\alpha) & = \sum_{\ell_1+\ell_2=\ell}E^{(\ell_1)}(\alpha)\cdot G^{(\ell_2)}(\alpha)                                                                   \\
                              & = \sum_{\ell_1+\ell_2=\ell}E_{\la \ell_1\ra}(\alpha)\cdot G_{g,w,\ell_2}(\alpha)\cdot g^{M-\ell_2}(\alpha)\cdot \lt(\psi\circ g(\alpha)\rt) \\
                              & = \psi\circ g(\alpha) \sum_{\ell_1+\ell_2=\ell}E_{\la \ell_1\ra}(\alpha)\cdot G_{g,w,\ell_2}(\alpha)\cdot g^{M-\ell_2}(\alpha)
\end{align*}
So define $$V_{\ell}(X)= \sum_{\ell_1+\ell_2=\ell}E_{\la \ell_1\ra}(X)\cdot G_{g,w,\ell_2}(X)\cdot g^{M-\ell_2}(X).$$ Then we have $(E\cdot G)^{(\ell)}(\alpha)=r(\alpha)\cdot V_{\ell}(\alpha)$. Since at error locations we have $E^{(i)}(\alpha)$ for all $0\leq i <M$ we have $V_{\ell}(\alpha)=0$. So this equality holds. Now $\deg(V_{\ell})\leq d\lt(\frac{q-1}2+M\rt)cq\eqqcolon D$ and $\deg(V_{\ell})\leq h+dM$. So we have $F$ to be a $r$-twisted  $(h+dM,M)$-pseudopolynomial of degree $D$. So if we find any $r$-twisted  $(h+dM,M)$-pseudopolynomial $F$ of degree $D$ via interpolation then the $E\cdot G$ and $F$, are both $r$-twisted. So by \TwistedTwolem there exists $A,B\in\bbF_q[X]$ such that $A\cdot (EG)=B\cdot F$. We can rewrite and say there exists $A,B\in\bbF_q[X]$ such that $A\cdot G=B\cdot F(X)$. So we have the following algorithm:

\begin{algorithm}
  \KwIn{Noisy received word, $r:\bbF_q\to \{0,1,-1\}$, with degree $d\leq O(\eps\sqrt{q})$ and error bound $e\leq \lt(\nicefrac18-\eps\rt)q$}
  \KwOut{Find the closest vector $\eta\circ f$ where $f\in\bbF_q[X]$ of degree at most $d$}
  \DontPrintSemicolon
  \caption{Decoding from $\nicefrac18-$th of distance of Quadratic Character Evaluation}
  \Begin{
    Set $M=\nicefrac{16}{\eps}\cdot d$, $c=\nicefrac{M}2$, $h=2\cdot e$, $D=d(\nicefrac{(q-1)}2+M)+cq$\;
    Use system of linear equations to find $r$-twisted $(h+dM,M)$-pseudopolynomial $F(X)$ of degree at most $D$. \;
    Complete factorization of $F(X)$ into distinct monic irreducible factors $\{h_1,\dots, h_k\}$ with $\mult(h_j,F)=k_j$. \;
    Compute the set $J=\lt\{j\in [t]\colon k_j\in \lt[\nicefrac{3q}8,\nicefrac{7q}8\rt]\rt\}$. Then compute $f(X)=\prod_{j\in J}h_j$. \;
    \Return{$f(X)$}
  }
\end{algorithm}
\subsubsection{Relating $F$, $G$ and Factor Multiplicities}
To make sense of the last step of separating out the high multiplicity factors we first have to look into $A,B$.


\paragraph*{Step 1: Relating $F$ and $G$}

We now show that any polynomial $F(X)$ obtained in the algorithm must satisfy a nontrivial algebraic relation with $G(X)$. Let $U_0,\dots, U_{M-1}$ be the polynomials such that $F^{(\ell)}(\alpha)=r(\alpha)\cdot U_{\ell}(\alpha)$ for all $0\leq \ell<M$, $\alpha\in\bbF_q$. So now we set parameters $t=\nicefrac38 M$ and $k=e+4dM$.
We claim that there exist nonzero polynomials $A(X)$ and $B(X)$ of the form
\[
  A(X)=\sum_{i=0}^{t-1} A_i(X)\Lambda^i(X),
  \qquad
  B(X)=\sum_{i=0}^{t+c-1} B_i(X)\Lambda^i(X),
\]
with 
\(
\deg(A_i)\le k\), \(
\deg(B_i)\le k+2e,
\)
such that
\(
A(X)F(X)=B(X)G(X).
\) We again argue by dimension counting. The total number of coefficients of the $A_i$ and $B_i$ equals 
\(
N=t(k+1)+(t+c)(k+2e+1).
\) Substituting the values of $t$ and $k$, we get
\begin{equation}\label{eq:FG-relate-num-var}
  N>
  \frac38 M(e+4dM)+\frac78 M(3e+4dM)
  =
  3Me+5dM^2.
\end{equation}
We regard these coefficients as variables and impose homogeneous linear conditions so that for every $0\le \ell<M$ and every $\alpha\in\mathbb F_q\setminus S$,
\begin{equation}\label{eq:twisted-relate-FG}
  (A\cdot F)^{(\ell)}(\alpha) = (B\cdot G)^{(\ell)}(\alpha).
\end{equation}
Consequently, the polynomial $A(X)F(X)-B(X)G(X)$ vanishes to multiplicity at least $M$ at every point of $\mathbb F_q\setminus S$. Hence it has at least $$M\cdot |\mathbb F_q\setminus S|\ge M(q-e)>
  \left(\frac78+\varepsilon\right)Mq$$
roots counted with multiplicity. On the other hand,
\[
  \deg(A\cdot F-B\cdot G)
  \le
  \max\{\deg(A\cdot F),\deg(B\cdot G)\}
  <
  \left(\frac78+\frac{\varepsilon}{2}\right)Mq.
\]
Thus, 
\(
A(X)F(X)=B(X)G(X).
\) It remains to show that Equation \eqref{eq:twisted-relate-FG} can indeed be enforced using fewer than $N$ linear constraints. For each $\ell$, define 
\(
A_{\langle \ell\rangle}(X)
=
\sum_{i=0}^{t-1}
(-1)^i A_i^{(\ell-i)}(X)\), \(
B_{\langle \ell\rangle}(X)
=
\sum_{i=0}^{t+c-1}
(-1)^i B_i^{(\ell-i)}(X).
\) By \DerivativeXqXlem we get 
\(
A^{(\ell)}(\alpha)=A^{\langle \ell\rangle}(\alpha)\) and \(
B^{(\ell)}(\alpha)=B^{\langle \ell\rangle}(\alpha)
\) for every $\alpha\in\mathbb F_q$. Now recall that 
\(
F^{(\ell)}(\alpha)=r(\alpha)U_\ell(\alpha).
\) Hence,
\begin{equation}\label{eq:derivative-of-AF}
  (A\cdot F)^{(\ell)}(\alpha)
  =
  \sum_{\ell_1+\ell_2=\ell}
  A^{(\ell_1)}(\alpha)F^{(\ell_2)}(\alpha)\implies
  (A\cdot F)^{(\ell)}(\alpha)
  =
  r(\alpha)\cdot
  \sum_{\ell_1+\ell_2=\ell}
  A_{\langle \ell_1\rangle}(\alpha)
  U_{\ell_2}(\alpha).
\end{equation}
Similarly, we get
\begin{equation}\label{eq:derivative-of-BG}
  G^{(\ell)}(\alpha)
  =
  (\chi\circ g)(\alpha)\,
  H_{g,w,\ell}(\alpha)\,
  g^{M-\ell}(\alpha),
  \implies
  (B\cdot G)^{(\ell)}(\alpha)
  =
  (\chi\circ g)(\alpha)\cdot
  \sum_{\ell_1+\ell_2=\ell}
  B_{\langle \ell_1\rangle}(\alpha)
  G_{g,w,\ell_2}(\alpha)
  g^{M-\ell_2}(\alpha).
\end{equation}

We therefore impose, for each $0\le \ell<M$, the polynomial identity
\begin{equation}\label{eq:solution-for-A-B}
  \sum_{\ell_1+\ell_2=\ell}
  A_{\langle \ell_1\rangle}(X)
  U_{\ell_2}(X)
  =
  \sum_{\ell_1+\ell_2=\ell}
  B_{\langle \ell_1\rangle}(X)
  G_{g,w,\ell_2}(X)
  g^{M-\ell_2}(X).
\end{equation}
Both sides have degree at most $3e+dM$. Hence, the total number of linear conditions arising from all these identities is at most \(
M(3e+dM+1),
\) which is strictly smaller than $N$ by Equation \eqref{eq:FG-relate-num-var}. Therefore there exists a nontrivial choice of $A(X),B(X)$ satisfying Equation \eqref{eq:solution-for-A-B}. Finally, combining Equations \eqref{eq:derivative-of-AF}, \eqref{eq:derivative-of-BG}, and \eqref{eq:solution-for-A-B}, and using the fact that $r(\alpha)=(\chi\circ g)(\alpha)$ for all $\alpha\in\mathbb F_q\setminus S$, we obtain Equation \eqref{eq:twisted-relate-FG}, and hence $A(X)F(X)=B(X)G(X)$.

\paragraph*{Step 2: Recovering the factors of $g(X)$}

We now use the identity $A(X)F(X)=B(X)G(X)$ to compare the irreducible factors of $F(X)$ and $G(X)$. The essential point is that the special form of $A(X)$ and $B(X)$ forces all their irreducible factors to occur with multiplicity very close to a multiple of $q$ by \PdegMultBoundlem. In contrast, the factors arising from $g(X)$ occur inside $G(X)$ with multiplicity approximately $q/2$ modulo $q$. Since $g(X)$ is squarefree, this allows us to identify precisely the irreducible factors of $g(X)$.

Now let $H(X)$ be an irreducible polynomial in $\mathbb F_q[X]$. Denote by
\(
  \mu_F,\mu_A,\mu_B,\mu_G,\mu_g
\) the multiplicities of $H(X)$ in $F(X),A(X),B(X),G(X),g(X)$ respectively. Since $g(X)$ is squarefree, 
\(
  \mu_g\in\{0,1\}.
\) From the identity
\(
  AF=BG,
\) we obtain
\(
  \mu_F=\mu_G+\mu_B-\mu_A.
\) 
Applying the lemma to $A$ and $B$, we have
\[
  \mu_B\in q\mathbb Z+[0,3e+4dM+\tfrac78 M]\quad\text{ and }\quad
  \mu_A\in q\mathbb Z+[0,e+4dM+\tfrac38 M].
\]
Therefore
\(
  \mu_F
  \in
  \mu_G+q\mathbb Z+[-e-5dM,\,3e+5dM].
\) Now
\(
  \mu_G=\left(\frac{q-1}{2}+M\right)\mu_g.
\) 

If $\mu_g=1$, then
\(
  \mu_G=\frac q2+\frac{M-1}{2},
\)
and hence
\(
  \mu_F
  \in
  q\mathbb Z+\frac q2+[-e-5dM,\,3e+6dM].
\) Using 
\( 
  e\le \left(\frac18-\varepsilon\right)q
\) 
and 
\(
  6dM<\varepsilon q,
\)
we obtain 
\(
  \mu_F\in q\mathbb Z+(3q/8,7q/8).
\) On the other hand, if $\mu_g=0$, then $\mu_G=0$, and thus
\[
  \mu_F
  \in
  q\mathbb Z+[-e-5dM,\,3e+5dM]
  \subseteq
  q\mathbb Z+(-q/8,3q/8).
\]

Hence, the set $J$ selected in Step~5 of the algorithm consists precisely of those irreducible factors which divide $g(X)$. Therefore, the polynomial produced in Step~5 is exactly $f(X)=g(X)$. 



\subsection{General Multiplicative Character of order \texorpdfstring{$m$}{m}}

Now we take the multiplicative character $\psi\in\Mq$or order $m>1$. Hence, we assume the polynomials are monic and don't have a factor of power $m$. Then by \WeilBoundPsiNondthm we have $$\lt|\sum_{\alpha}\psi((g_1\cdot g_2)(\alpha))\rt|< O(m\sqrt{q})$$Therefore the hamming distance is at least $\lt(1-\frac1m\rt)q$ which gives the distance of the code. So the unique decoding radius is $\frac12\lt(1-\frac1m\rt)q$.

Now we can also think of the character as $\psi:\bbF_q\to \{0\}\cup\bbF_q^{*^{(\nicefrac{(q-1)}m)}}$ where $\alpha\mapsto \alpha^{\nicefrac{(q-1)}m}$. So here we will consider the multiplicative as an map from $\bbF_q\to \{0\}\cup\bbF_q^{*^{(\nicefrac{(q-1)}m)}}$. Like for the case of quadratic character we will not prove the correctness of the algorithm instead we just give the pseudocode of the algorithm below. The details of the correctness of the algorithm is given in \cite{Kopparty2026}.


\begin{algorithm}
  \KwIn{Noisy received word, $r:\bbF_q\to \{0,1,-1\}$, with degree $d\leq O(\frac{\eps}{m}\sqrt{q})$ and error bound $e\leq \lt(\nicefrac1{12}-\eps\rt)q$}
  \KwOut{Find the closest vector $\psi\circ f$ where $f\in\bbF_q[X]$ of degree at most $d$}
  \DontPrintSemicolon
  \caption{Decoding from $\nicefrac1{12}-$th of distance of $m^{th}$-order Character Evaluation}
  \Begin{
    Set $M=\nicefrac{16}{\eps}\cdot dm$, $c=\nicefrac{M}2$, $h=2\cdot e$, $D=d((q-1)\cdot \nicefrac{(m-1)}{m}+M)+cq$\; 
    \For{$k\in[m-1]$}{
      Use system of linear equations to find $r^k$-twisted $(h+(m-1)dM,M)$-pseudopolynomial $F_k(X)$ of degree at most $D$. \;
      Complete factorization of $F(X)$ into distinct monic irreducible factors $\{h_{1,k},\dots, h_{t_k,k}\}$ with $\mult(h_{j,k},F_k)=d_{j,k}$. 
    }
    Let the irreducible factors are $H_1,\dots, H_t$ with $\mult(H_j,F_k)=d_{j,k}$.
    \For{$j\in [t]$}{
      Let $\mu_j\in\{0,1,\dots, m-1\}$ be the unique number such the for all $k\in[m-1]$ $$d_{k,j}-\frac{k\cdot \mu_j}{m}\cdot (q-1)\in\lt(-\lt(\frac1{12}-\eps\rt)q,\lt(\frac3{12}-\eps\rt)q  \rt)$$
    }
    Compute $f(X)=\prod_{j=1}^th_j^{\mu_j}(X)$.\;
    \Return{$f(X)$}
  }
\end{algorithm}