\chapter{Exponential Sums over Finite Fields}

An \emph{exponential sum} over $\bbF_q$ is a finite sum of the form $\sum_{\alpha\in\bbF_q} f(\alpha)$, where $f$ takes values on the unit circle in $\bbC$. Because $\bbF_q$ carries two group structures simultaneously --- the additive group $(\bbF_q,+)$ and the multiplicative group $(\bbF_q^*,\times)$ --- there are two natural families of unit-circle-valued homomorphisms, called \emph{additive} and \emph{multiplicative characters}, and exponential sums arise by combining them in various ways. These sums encode deep arithmetic about $\bbF_q$: the absolute value $|\sum_\alpha f(\alpha)|$ measures how uniformly distributed the values of $f$ are, and proving that it is small (typically $O(\sqrt{q})$) has direct consequences for counting solutions to polynomial equations, the distribution of quadratic residues, and the structure of Gauss and Kloosterman sums.


\paragraph*{Two Types of Characters.}
A finite field $\bbF_q$ has two natural families of characters. \emph{Additive characters} $\chi:\bbF_q\to S^1$ are group homomorphisms from $(\bbF_q,+)$; every one is of the form $\chi_\beta(\alpha)=e^{2\pi i\,\tr(\beta\alpha)/p}$ for some $\beta\in\bbF_q$, where $\tr:\bbF_q\to\bbF_p$ is the absolute trace. \emph{Multiplicative characters} $\psi:\bbF_q^*\to S^1$ are group homomorphisms from the cyclic group $(\bbF_q^*,\times)$ of order $q-1$; a particularly important special case is the \emph{quadratic character} $\eta$, the finite-field analog of the Legendre symbol. The interplay between these two families --- their orthogonality, their products, and their behavior under field extensions --- is the engine driving all the results in this chapter.

\paragraph*{Chapter Outline.}
The chapter is organized in three layers: (1) the algebraic foundations of characters, (2) the core classes of exponential sums and their evaluation, and (3) the $L$-function machinery that lifts these evaluations to field extensions.

\S \Autoref{sec:chars-finite-abelian}--\S \Autoref{sec:chars-finite-fields}: \emph{Characters of Finite Abelian Groups and Finite Fields} lay the foundations. We first define characters abstractly as homomorphisms $G\to S^1$ for a finite abelian group $G$, establish the dual group $\hG\cong G$, and prove the orthogonality relations. We then specialize to $\bbF_q$: the canonical additive character $\chi_1$ is constructed via the trace map, and it is shown that all additive (resp.\ multiplicative) characters are twists of $\chi_1$ (resp.\ a generator of $\Mq$). Both character groups satisfy orthogonality: $\sum_\alpha\chi(\alpha)=0$ for $\chi\neq\chi_0$, and $\sum_\alpha\psi(\alpha)=0$ for $\psi\neq\psi_0$.

\S \Autoref{sec:gaussian-sums}: \emph{Gaussian Sums} are the first hybrid objects: $G(\psi,\chi)=\sum_{\alpha\in\bbF_q^*}\psi(\alpha)\chi(\alpha)$. They bridge the additive and multiplicative structures and are evaluated in closed form. The central result is that when both characters are non-trivial, $|G(\psi,\chi)|=\sqrt{q}$ --- a non-trivial cancellation that lies at the heart of all subsequent $O(\sqrt{q})$ bounds. Gaussian sums also satisfy a key inversion identity, expressing $\psi$ as an additive transform of itself via $G$.

\S \Autoref{sec:jacobi-sums}: \emph{Jacobi Sums} are purely multiplicative sums: $J(\lm_1,\dots,\lm_k)=\sum_{\alpha_1+\cdots+\alpha_k=1}\lm_1(\alpha_1)\cdots\lm_k(\alpha_k)$. They arise naturally in counting solutions to additive equations with multiplicative weights and are connected to Gaussian sums by the fundamental relation $J(\lm_1,\lm_2)=G(\lm_1,\chi)G(\lm_2,\chi)/G(\lm_1\lm_2,\chi)$ when $\lm_1\lm_2\neq\psi_0$. Via this relation, $|J(\lm_1,\lm_2)|=\sqrt{q}$ when all characters are non-trivial.

\S \Autoref{sec:l-functions}: \emph{$L$-functions} build the analytic infrastructure. The Riemann zeta function $\zeta(s)$ and its function-field analogue over $\bbF_q[X]$ are introduced, emphasising the parallel: just as $\zeta(s)$ factors as an Euler product over primes of $\bbZ$, the zeta function of $\bbF_q[X]$ factors over monic irreducible polynomials, and twisting by a multiplicative character $\chi$ yields $L(s,\chi)$. The key theorem is that $\ovL(z,\chi)$ is a polynomial of finite degree in $z=q^{-s}$, whose reciprocal roots carry the arithmetic information, and the Riemann Hypothesis for $\bbF_q[X]$ (a theorem) asserts all roots have absolute value $q^{-1/2}$.

\S \Autoref{sec:lifting-gauss-jacobi}: \emph{Lifting of Gaussian and Jacobi Sums} uses the $L$-function machinery to relate sums over $\bbF_q$ with sums over extensions $\bbF_{q^r}$. The \emph{Davenport--Hasse theorem} is the central result: if $\psi,\chi$ are lifted to $\psi^{(r)},\chi^{(r)}$ on $\bbF_{q^r}$, then $G(\psi^{(r)},\chi^{(r)})=(-1)^{r-1}G(\psi,\chi)^r$. The analogous formula for Jacobi sums follows. These lifting formulas are indispensable for the proof of the Weil bounds in the following chapter.

\paragraph*{Further Sum Families.}
The final four sections study exponential sums that arise in applications to classical problems.

\S \Autoref{sec:quadratic-char}: \emph{Character Sums via the Quadratic Character} yields an explicit evaluation of $G(\eta,\chi_1)$ in terms of congruence conditions on $p$: the result is $(-1)^{r-1}i^{(p-1)^2r/4}q^{1/2}$, a finite-field version of Gauss's famous determination of the sign of the classical Gauss sum.

\S \Autoref{sec:jacobsthal}: \emph{Jacobsthal Sums} $H_n(\alpha)=\sum_\gamma\eta(\gamma^{n+1}+\alpha\gamma)$ mix the quadratic character with a polynomial twist. Introduced by Jacobsthal in 1907 to study representations of primes by quadratic forms, they satisfy a duplication recursion $I_{2n}=I_n+H_n$ and their connection to Jacobi sums gives explicit evaluations.

\S \Autoref{sec:kloosterman}: \emph{Kloosterman Sums} $K(\chi;a,b)=\sum_{\gamma\in\bbF_q^*}\chi(a\gamma+b\gamma^{-1})$ twist a single additive character with the inversion map. Introduced by Kloosterman in 1926 in the context of the circle method for quaternary quadratic forms, they have no closed-form evaluation, but the Weil bound gives $|K(\chi;a,b)|\leq 2\sqrt{q}$ for $ab\neq 0$.

\S \Autoref{sec:salie}: \emph{Sali\'{e} Sums} $S(\chi,\tau)=\sum_\gamma\eta(\gamma)\chi(\gamma)\tau(\gamma^{-1})$ insert the quadratic character $\eta$ as a weight into a Kloosterman-type sum. This mild modification makes exact evaluation possible, yielding $S(\chi,\tau)=G(\eta,\chi)\sum_{b^2=4a}\chi(b)$, in striking contrast with the intractability of bare Kloosterman sums.

\input{chapters/character_sums/characters_of_finite_abelian_groups}
\input{chapters/character_sums/characters_of_finite_fields}
\input{chapters/character_sums/gaussian_sums}
\input{chapters/character_sums/jacobi_sums}
\input{chapters/character_sums/l_functions}
\input{chapters/character_sums/lifting-gauss-jacobi}
\input{chapters/character_sums/kloosterman}
\input{chapters/character_sums/quadratic-char}
\input{chapters/character_sums/jacobsthal}
\input{chapters/character_sums/salie}
