\chapter{Equations over Finite Fields}

The central question of this chapter is: given a polynomial $f\in\bbF_q[X_1,\dots,X_n]$, how many solutions does $f(X_1,\dots,X_n)=0$ have in $\bbF_q^n$? We write $\mcV(f)$ for the solution set and $Z(f)=|\mcV(f)|$ for its size; for a system $f_1=\cdots=f_k=0$ we write $\mcV(f_1,\dots,f_k)=\mcV(f_1)\cap\cdots\cap\mcV(f_k)$ and $Z(f_1,\dots,f_k)$ for the common solution count. Variants $\ov{\mcV}$ and $\ov{Z}$ track the non-trivial (nonzero) solutions. The notation mirrors that of varieties of algebraic geometry: $\mcV(f)$ is the affine variety cut out by $f$.

The chapter develops two complementary strategies. The first is \emph{direct combinatorics}: for special polynomial families (univariate polynomials, systems of low total degree, quadratic forms, diagonal equations) one can write exact formulas or prove sharp divisibility constraints on $Z(f)$. The second is \emph{character sum methods}: the indicator of whether $\alpha$ is a solution can be encoded as an additive character sum, turning the problem of counting solutions into the problem of bounding $\sum_\alpha\chi(f(\alpha))$. These two strategies reinforce each other throughout the chapter.

\paragraph*{Chapter Outline.}

\S \Autoref{sec:elementary-results}: \emph{Some Elementary Results} establishes the foundational exact and divisibility results. For univariate polynomials the \emph{K\"{o}nig--Rados theorem} identifies $\ov{Z}(f)=q-1-\Rank(M_f)$, where $M_f$ is the left circulant matrix of $f$'s coefficients. For multivariate systems the \emph{Chevalley--Warning theorem} proves that if the total degree of a system is less than the number of variables then $p\mid Z(f_1,\dots,f_k)$; in particular, a system with no constant term has a non-trivial solution. Warning's strengthening bounds the number of solutions from below by $q^{n-\deg}$.

\S \Autoref{sec:character-sum-poly}: \emph{Character Sums with Polynomials} introduces the \emph{Weil sum} $W(\chi;f)=\sum_{\alpha\in\bbF_q}\chi(f(\alpha))$ and its multiplicative and mixed variants $W(\psi;f)$ and $W(\psi,\chi;f,g)$. Exact evaluations are carried out for shifted monomials $\chi(ac^n+b)$ by expressing the $d$-th power map in terms of multiplicative characters of order $d=\gcd(n,q-1)$ and reducing to Gaussian sums. These exact formulas serve as the model for all subsequent character sum estimates.

\S \Autoref{sec:average-solutions}: \emph{Average Number of Solutions} studies the distribution of $Z(f)$ over the space $\Om_d=\bbF_q^{\leq d}[X_1,\dots,X_n]$ of all polynomials of degree $\leq d$. The average is $q^{n-1}$ --- a random polynomial equation in $n$ variables has $q^{n-1}$ solutions on average --- and the variance is $q^{n-1}-q^{n-2}$. These results show that most polynomials have close to the expected number of solutions, and set the scale for the deviations studied in later chapters.

\S \Autoref{sec:quadratic-forms}: \emph{Quadratic Forms} specialises to degree-$2$ homogeneous polynomials. A quadratic form $f=X^\top C_f X$ is represented by its symmetric coefficient matrix $C_f$, and two forms are equivalent under non-singular linear substitution iff their coefficient matrices are congruent. The number of solutions to $f(X_1,\dots,X_n)=\alpha$ is computed explicitly using the $\theta$-function convolution, and depends only on $n$, the rank of $C_f$, and whether $\alpha=0$ or not. The character sum machinery of the previous section applies here too: $Z(f=\alpha)$ is expressed as a sum of Gaussian sums weighted by $\eta(\det C_f)$.

\S \Autoref{sec:diagonal-equations}: \emph{Diagonal Equations} treats equations of the form $a_1X_1^{k_1}+\cdots+a_nX_n^{k_n}=\alpha$. The solution count factors as a sum over $\alpha_1+\cdots+\alpha_n=\alpha$ of products of single-variable counts $Z(X_i^{k_i}=a_i^{-1}\alpha_i)$, each of which is a sum over characters of order $d_i=\gcd(k_i,q-1)$. Collecting terms via the Jacobi sum formula gives $$Z\!\left(a_1X_1^{k_1}+\cdots+a_nX_n^{k_n}=\alpha\right)=q^{n-1}+\sum_{t_1=1}^{d_1-1}\cdots\sum_{t_n=1}^{d_n-1}\bar{\lambda}_1^{t_1}(a_1)\cdots\bar{\lambda}_n^{t_n}(a_n)\cdot J_\alpha(\lambda_1^{t_1},\dots,\lambda_n^{t_n}),$$ and the bound $|J|=q^{(n-1)/2}$ yields $|Z-q^{n-1}|\leq (q-1)q^{n/2-1}\prod_i(d_i-1)$. 
\input{chapters/equations-over-fq/elementary-results}
\input{chapters/equations-over-fq/character-sum-of-poly}
\input{chapters/equations-over-fq/average-solutions}
\input{chapters/equations-over-fq/quadratic-forms}
\input{chapters/equations-over-fq/diagonal-equations}
